---
title: "homework2"
author: "KiseokUchicago"
date: "2020-10-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Homework2
## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

### (2)
The built-in Matlab command to construct the LU-factorization of a given matrix A is
[L,U]=lu(A) (the built-in command in Octave is the same, in R it is contained in the Ma-
trix library (loaded with library(Matrix)) and obtained by writing expand(lu(A)). Use this
command to find the factorization A = LU.

### (2)-(a)

```{r}
library(Matrix)

A = matrix( c(2,-1,0,-1,2, -1, 0, -1, 2), nrow=3, ncol=3)
A

LU_A <- expand(lu(A))
LU_A

# L is
LU_A$L

# U is
LU_A$U
LU_A$L %*% LU_A$U # equals A

```

### (2)-(b)
Find the Matlab or R (or Octave or Julia) built-in command to solve the systems of the form
Ax = b

solve() function works to solve the Ax=b

```{r}
A = matrix( c(2,-1,0,-1,2, -1, 0, -1, 2), nrow=3, ncol=3)
b = c((pi^2)/4,0,-((pi^2)/4))
solve(A,b)

```

### (2)-(c)
Write a program implementing Gaussian elimination, including row swaps.

```{r}
# We will write a program that works for 3 by 3 matrix

A = matrix( c(2,-1,0,-1,2, -1, 0, -1, 2), nrow=3, ncol=3)
b = c((pi^2)/4,0,-((pi^2)/4))

# make augmented matrix
m <- cbind(A,b)

# recursive function for row exchange and elimination
recursive_f <- function(m,col_n){
  i <- col_n # col_n means column number
  j <- 1
  a = c(1:dim(m)[1])
  while(m[i,i]==0){ # if (i,i) is zero, exchange rows with next row until there is non-zero pivot
    m[c(i,i+a[j]), ]  <- m[c(i+a[j],i), ]
    print(paste0('row exchange ',i, 'with ',i+a[j] ))
    print(m)
    j <- j+1
    if(m[i,i]!=0 | j == dim(m)[1]){ # break out of loop if we have non-zero pivot, or there is all rows have 0 in the column i.
      break
    }  
  } 
  # subtract the rows below the pivot to and make it zero
  for (k in (i+1):(dim(m)[1])){
    # print(k)}
    print (paste0('subtract row',k,' with row',i))
    m[k,] <- m[k,] - (m[k,i]/m[i,i])*m[i,]
    print(m)
  }
  return(m)
}

# recursive function for backward substitution (Diagonalizing U for convenience)
backsub_recursive <- function(m,col_n){
  g <- col_n
  if (m[g,g]==0){
    print(paste0('no or many solution for x',g))
    return(m)
  } else{
    for (h in g:2){
      # print(g)}
      print (paste0('subtract row',h-1,' with row',g))
      m[h-1,] <- m[h-1,] - (m[h-1,g]/m[g,g])*m[g,]
      print(m)
      return(m)
    }
  } 
}

# solve the equation Ax=b
A = matrix( c(2,-1,0,-1,2, -1, 0, -1, 2), nrow=3, ncol=3)
b = c((pi^2)/4,0,-((pi^2)/4))

# make augmented matrix
m <- cbind(A,b)

for (l in 1:(dim(m)[1]-1)){
  # print(l)}
  m <- recursive_f(m,l)
  print(paste0('done with pivot',l))
}

# this is Upper triangular matrix
U <- m
U

# back substitution
# We will make U diagonal for back substitution
for (g in dim(m)[1]:2){
  U <- backsub_recursive(U,g)
}

# get the solution x vector
x =c()
for (i in 1:dim(U)[1]){
  x[i]=U[i,dim(U)[2]] / U[i,i]
}

x   # by using round off values of pi, we did not get 0 in x2
round(x,10) 

```





