---
title: "Homework1"
author: "KiseokUchicago"
date: "2020-10-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```
## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

(2) Markov Matrice problem: This is a computational exercise, meant to be completed using Matlab, Octave, R, or Julia, related to methods for solving linear systems by iterative techniques (that is, constructing a sequence of approximate vectors which \converges" to the desired solution).

### (2)-(a)

```{r}
u0 = c(1,0)
A = matrix( c(.2,.8,.7,.3), nrow=2, ncol=2)
A
u1 = A%*%u0
u2 = A%*%u1
u3 = A%*%u2
u4 = A%*%u3
```
The common property of u1, u2, u3, u4 is that the entries are summed to 1.

### (2)-(b)
```{r}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(plotly)
library(htmlwidgets)

```


```{r}
u5 = A%*%u4
u6 = A%*%u5
u7 = A%*%u6

df_u <- t(cbind(u0, u1, u2, u3, u4, u5, u6, u7))
colnames(df_u) <- c('x','y')
df_u <- data.frame(df_u)
rownames(df_u) <- c('u0', 'u1', 'u2', 'u3', 'u4', 'u5', 'u6', 'u7')
df_u$label <- c('0', '1', '2', '3', '4', '5', '6', '7')
df_u

p1 <- ggplot(df_u,aes_string(x = "x",y = "y")) + geom_point(color='maroon') + theme_cowplot() + geom_text_repel(aes(label=label))
p1

```

### (2)-(c)

```{r}
u=c(1,0)
for (i in 1:1000000){
  u <- A %*% u
}

u_star <- u

u_star - A %*% u_star

```







