---
title: "Homework6"
author: "KiseokUchicago"
date: "2020-11-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```
## Homework6
## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

### (4)-(a)
(The "power method" for computing eigenvalues) Suppose that a real n x n matrix A has a set of
n linearly independent eigenvectors.

```{r}
library(expm)
A1 = matrix( c(0.7, 0.1, 0.2,
               0.3,0.5,0.2,
               0.2, 0.7, 0.1), nrow=3, ncol=3)

A2 = matrix( c(0.90, 0.05, 0.05,
               0.10,0.80,0.10,
               0.01, 0.01, 0.98), nrow=3, ncol=3)

euc_norm <- function(x) sqrt(sum(x^2))

u0= c(1,0,0)

df_uk <- data.frame(k =double(), ukt_length_A1=double(), ukt_length_A2=double())

for (k in c(2, 4, 8, 16, 100000000)){
  print(k)
  uk_A1 <- (A1 %^% k)%*%u0
  uk_A1_div <- (uk_A1)/euc_norm(uk_A1)
  print( (uk_A1)/euc_norm(uk_A1))
  
  uk_A2 <- (A2 %^% k)%*%u0
  uk_A2_div <- (uk_A2)/euc_norm(uk_A2)
  print( (uk_A2)/euc_norm(uk_A2))  
  
  df_uk <- rbind(df_uk, c(k, round(uk_A1_div,4), round(uk_A2_div,4 )))
  
}

colnames(df_uk) <- c('k', 'ukT/||uk||', 'from','uk=A1k*u0',' ukT/||uk||', 'from ','uk=A2k*u0' )

df_uk
```

### (4)-(b)
For which matrix do we find faster convergence of uTk =|uk| to the true unit eigenvector? Explain why. (Hint: Think about how you might try to prove (1).)

```{r}
library(tictoc)

tic.clearlog()

for (k in c(2, 4, 8, 16, 10000000000000000)){
  tic()
  uk_A1 <- (A1 %^% k)%*%u0
  uk_A1_div <- (uk_A1)/euc_norm(uk_A1)
  toc(log=T)
}


for (k in c(2, 4, 8, 16, 10000000000000000)){
  tic()
  uk_A2 <- (A2 %^% k)%*%u0
  uk_A2_div <- (uk_A2)/euc_norm(uk_A2)
  toc(log=T)
}

log.lst <- tic.log(format = FALSE)
tic.clearlog()

timings <- unlist(lapply(log.lst, function(x) x$toc - x$tic))
timings
```
A1 has faster convergence time to unit eigenvector compared to that of A2. 


