---
title: "Homework5"
author: "KiseokUchicago"
date: "2020-11-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```
## Homework5
## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

### (2)-(a)
Give the R, Matlab, Octave, or Julia commands to produce a QR decomposition for the matrices in Questions 4(b) and 4(c) in the joint Stat 24300/30750 portion of this assignment.

```{r '2-a'}
library(matlib)

# A from question 4.
A = matrix( c(1, 1, -2, 1, -1, 4), nrow=3, ncol=2)
b = c(1, 2, 7)

# QR decompostion of A
qr_A <- qr(A)
Q <- qr.Q(qr_A)
Q
R <- qr.R(qr_A)
R
qr.Q(qr_A) %*% qr.R(qr_A) # brings back A. decomposition success.

```
I think the question have issue referring to the right matrix.

### (2)-(b)
The question (4)-(a) does not have {q1, q2, q3}. Something has to change in order for the  problem to make sense.

### (3) Watkins Exercise 3.2.8
Show that Q is orthogonal, then ||Q||2=1, ||Q-1||2=1, and k2(Q)=1

```{r}
random_A = matrix( runif(9), nrow=3, ncol=3)
random_QR <- qr(random_A)
random_Q <- qr.Q(random_QR)

base::norm(random_Q, type='2')
base::norm(inv(random_Q), type='2')

# kappa_2 is the value multiplying thses two values above.
kappa_A = base::norm(random_Q, type='2') * base::norm(inv(random_Q), type='2')
kappa_A
```

### (4) Watkins Exercise 3.2.12
QTx = [y, 0]

### (4)-(a) 
Give the geometric interpretation of this fact.

This satisfies if and only if x * sin(theta) = y * cos(theta).
Therefore, tan(theta) = x/y. This always holds true.
Geometrically, QTx = [y, 0] means that QT matrix is an linear transformation where it transforms any coordinate x=(x1, x2) in to (sqrt(xTx), 0), which is a point on the y-axis intersecting the circle x is in.

### (4)-(b) 
Show that cos(theta)=x1/sqrt(x1^2+x2^2), sin(theta)=x2/sqrt(x1^2+x2^2), then y = ||x||2

y = cos(theta)x1+sin(theta)x2
When we substitute cos and sin, then we get y = x1^2/sqrt(x1^2+x2^2) + x2^2/sqrt(x1^2+x2^2) = sqrt(x1^2+x2^2) = ||x||2

### (5) Watkins Exercise 3.4.33
Implementation of Gram-Schmidt 

```{r}
# classical Gram-Schmidt algorithm

euc_norm <- function(x) sqrt(sum(x^2))

classic_GS <- function(A){
  r = matrix(nrow=dim(A)[1], ncol=dim(A)[2])
  for (k in 1:dim(A)[2]){
    if(euc_norm(A[,k])==0){
      print('Length of column is 0. End of function.')
      stop()
    } # break out of loop if length is 0
    k_vec=1:k
    for (i in k_vec[-k]){
      r[i,k] <- sum(A[,i]*A[,k])
      A[,k] <- A[,k] - r[i,k]*A[,i]
    }
    A[,k] <- A[,k]/(euc_norm(A[,k])) 
  } 
  return(A)
}

# Original matrix
A = matrix( runif(9, 1, 10), nrow=3, ncol=3)
A

# Q matrix after Gram-Schmidt 
classic_GS(A)

# check all column being orthonormal
CGS_A <- classic_GS(A)

for (k in 1:dim(CGS_A)[2]){
  print(euc_norm(CGS_A[,k]))
} # success

# checking if flag is working
A = matrix( c(0,0,0,1,2,3,4,5,6), nrow=3, ncol=3)
classic_GS(A)

```


```{r}
# Modified Gram-Schmidt algorithm

modified_GS <- function(A){
  for (k in 1:dim(A)[2]){
    if(euc_norm(A[,k])==0){
      print('Length of column is 0. End of function.')
      stop()
    } # break out of loop if length is 0
    k_vec=1:k
    for (i in k_vec[-k]){
      r_ik <- sum(A[,i]*A[,k])
      A[,k] <- A[,k] - r_ik*A[,i]
    }
    A[,k] <- A[,k]/(euc_norm(A[,k])) 
  } 
  return(A)
}

# Original matrix
A = matrix( runif(9, 1, 10), nrow=3, ncol=3)
A

# Q matrix after Gram-Schmidt 
modified_GS(A)

# check all column being orthonormal
MGS_A <- modified_GS(A)

for (k in 1:dim(CGS_A)[2]){
  print(euc_norm(CGS_A[,k]))
} # success

# checking if flag is working
A = matrix( c(0,0,0,1,2,3,4,5,6), nrow=3, ncol=3)
modified_GS(A)

```

The results for two algorithms are the same.

```{r}
A = matrix( runif(9, 1, 10), nrow=3, ncol=3)
classic_GS(A) - modified_GS(A)
```







