---
title: "Homework1"
author: "KiseokUchicago"
date: "2020-10-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

### (2)
Markov Matrice problem: This is a computational exercise, meant to be completed using Matlab, Octave, R, or Julia, related to methods for solving linear systems by iterative techniques (that is, constructing a sequence of approximate vectors which \converges" to the desired solution).

### (2)-(a)

```{r}
u0 = c(1,0)
A = matrix( c(.2,.8,.7,.3), nrow=2, ncol=2)
A
u1 = A%*%u0
u2 = A%*%u1
u3 = A%*%u2
u4 = A%*%u3
```
The common property of u1, u2, u3, u4 is that the entries are summed to 1.

### (2)-(b)
```{r}
library(ggplot2)
library(cowplot)
library(ggrepel)
library(plotly)
library(htmlwidgets)

```


```{r}
u5 = A%*%u4
u6 = A%*%u5
u7 = A%*%u6

df_u <- t(cbind(u0, u1, u2, u3, u4, u5, u6, u7))
colnames(df_u) <- c('x','y')
df_u <- data.frame(df_u)
rownames(df_u) <- c('u0', 'u1', 'u2', 'u3', 'u4', 'u5', 'u6', 'u7')
df_u$label <- c('0', '1', '2', '3', '4', '5', '6', '7')
df_u

p1 <- ggplot(df_u,aes_string(x = "x",y = "y")) + geom_point(color='maroon') + theme_cowplot() + geom_text_repel(aes(label=label))
p1

```

### (2)-(c)

```{r}
u=c(1,0)
for (i in 1:1000000){
  u <- A %*% u
}

u_star <- u
u_star

u_star - A %*% u_star
```
Therefore, Au_star is approximately u_star, given that 1000000 is a big number. When tested with bigger numbers, the results stays the same. Thus, it could be said that the approximation is verifiable.

### (4)
Let n >= 1 be given. For each vector v = (v1; : : : ; vn) consisting of an n-tuple of real numbers, let f(v) = max(vi : i = 1,...,n) be the largest entry in v.


```{r}
# function for finding maximum value and its indices

max_finder <- function(v){
  if(length(v)==0){ # if there is no element in the vector, show error message
    print('please provide a vector with at least 1 element')
  } else {
    # get the maximum value by comparing 2 elements at once (sequentially)
    max=v[1]
    for (i in 1:length(v)){
      if (max < v[i]){
        max = v[i]
      }
    }
    # add a function to get the indices with max value
    index_list=c()
    for (i in 1:length(v)){
      if (v[i]== max){
        index_list <- c(index_list, i) 
      }
    }
    print('The indices with the maximum value are ')
    print(index_list)
    print('The maximum value is')
    return(max)
  }
}
```

```{r}
# test 1: getting the proper error message
v=c()
max_finder(v)

# test 2: getting the proper maximum value
v1 = c(1, 3, 4, 9, 5, 6)
max_finder(v1)

# test 3: getting the proper indices for ties in maximum
v2 = c(1, 3, 4, 9, 5, 6, 9, 9)
max_finder(v2)        
  
```


The next coding is just done for additional applications when working with multiple vectors(not asked from homework)
```{r}
# (Appendix) when working with a matrix of vectors
v1 = c(1, 3, 4, 9, 5, 6)
v2 = c(5, 10, 12, 11, 16, 8)
v3 = c(2, 2, 2, 2, 2, 19)

v_mat <- rbind(v1, v2, v3)
dim(v_mat)

max=v_mat[1,1]
for (i in 1:dim(v_mat)[1]){
  for (j in 1:dim(v_mat)[2]){
    if ((max < v_mat[i,j]) == TRUE){
      max = v_mat[i,j]
      print(max)
    } 
  }
}
max
```




