---
title: "Homework7"
author: "KiseokUchicago"
date: "2020-12-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```
## Homework7
## Written assignment for Stat 30750
Professor: **Fatma Terzioglu**  
Student: **Kiseok Lee**  

### (2) Singular value decomposition
### (2)-(a)

```{r 'a'}
library(dplyr)

output=matrix(NA, nrow=300, ncol=20)

for (i in 1:300){
  A <- matrix(nrow=40, ncol=20)
  for (k in 1:dim(A)[1]){
      A[k,]=rnorm(dim(A)[2], mean=0, sd=1)
  }
  dim(A)
  (s <- svd(A))
  dim(s$u)
  dim(s$v)
  # D <- diag(s$d)
  # s$u %*% D %*% t(s$v) #  X = U D V'
  # t(s$u) %*% X %*% s$v #  D = U' X V 
  output[i,] <- s$d
}

dim(output)
output_sqr <- sqrt(output)
df_output_sig <- as.data.frame(output_sqr)

head(df_output_sig)

```

### (2)-(b)
Plot a histogram of the largest singular value from the 300 matrices.

```{r 'b'}
hist(df_output_sig$V1, breaks=15, col="darkblue", border="slateblue", main="Histogram of largest sv", xlab = 'sigma_1')

```

### (2)-(c)
Plot a histogram of the smallest singular value from the 300 matrices.

```{r 'c'}
hist(df_output_sig$V20, breaks=15, col="darkblue", border="slateblue", main="Histogram of smallest sv", xlab = 'sigma_20')

```

### (2)-(d)
Plot a curve of the mean of the 20 singular values obtained from the 300 matrices,
using the eigenvalue order(from largest to smallest) as the horizontal axis.

```{r 'd'}
library(ggplot2)
table <- sapply(df_output_sig, mean, 1)
df_mean <- data.frame(mean=table)
rownames(df_mean) <- 1:20
df_mean2 <- tibble::rownames_to_column(df_mean, 'sigma')

plot(df_mean2, type ='l', main='Mean of 20 singular values')

```

### (2)-(e)
Rank of A is equal to number of sigma, 20. Since A=USVt, A is similar to S (A ~ S, S: singular value diagonal matrix). This means that the eigenvalue of two matrices are the same. Therefore, the rank, which is equal to the number of non-zero eigenvalues, of A and S is same.

